Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores with Time Series Modeling

### 2020-04-14 Ben_Manuscripts/stochastic_transport/Final_Draft_JPCB.tex
Minor additional comments
Update draft.

  Up through our previous studies, our molecular models have provided valuable
  qualitative mechanistic insight with some quantitative support. This insight
  already allows us to speculate about new LLC monomer designs. However, they
  would be of greater value to a larger set of researchers if we could provide
  quantitative predictions of macroscopic observables such as solute flux and
  selectivity. Due to the size of the types of systems we are studying (at
  least 62,000 atoms) it is prohibitively expensive and time-consuming to run
  simulations longer than those performed for this work ($\sim$5 $\mu$s).  Even
  with the relatively large size of our system, we only simulate 24 solute
  trajectories per unit cell in order to minimize solute-solute interactions.
  This results in relatively high uncertainties in observables such as mean
  squared displacement (MSD), preventing reliable long timescale predictions.
  We are in need of a way of studying the collective motion of a much larger
  set of solutes over much longer timescales.

  Mathematical descriptions of transport in complex separations membranes are a
  powerful way to understand mechanisms and formulate design principles.
  \cite{vinh-thang_predictive_2013,geens_transport_2006,darvishmanesh_mass_2016}
  The complexity of a well-fit model generally parallels the complexity of the
  transport mechanism being studied, as well as the transport information the
  model conveys. In dense homogeneous membranes, the solution-diffusion model
  can extract diffusion and partition coefficients and has successfully
  predicted solute transport rates.~\cite{wijmans_solution-diffusion_1995}
  Analogously, pore-flow models yield predictions of diffusion coefficients and
  solute transport rates in nanoporous membranes.~\cite{paul_diffusive_1974}
  Modern single particle tracking approaches have taken researchers beyond
  continuum modeling allowing them to characterize complex diffusive
  behavior.~\cite{manzo_review_2015} At the molecular level, one can use
  molecular dynamics (MD) simulations to study both single particle dynamics
  and bulk transport properties with atomic-level
  insight.~\cite{coscia_chemically_2019,maginn_best_2018} All of these
  approaches facilitate generation of hypotheses about the molecular origins of
  separations by attempting to give a more intuitive understanding of how
  solutes move as a function of their environment, in turn suggesting
  experiments that could be performed.

@@

  %BJC: replacement to above.
  We constructed our first model in order to characterize the hopping and
  trapping behavior displayed by solutes in our MD simulations. This type 
  of non-Brownian dynamical behavior results in subdiffusive transport which
  is well-described by anomalous diffusion (AD) theory.~\cite{metzler_random_2000,bouchaud_anomalous_1990,klages_anomalous_2008}

@@

@@

  Despite relatively good predictions of the MD MSD, qualitative mismatch
  between simulated MSDDM and MD trajectories suggest that the MSDDM may be
  getting the right answers for the wrong reasons. We plotted typical
  realizations of the MSDDM for each solute and compared them to MD in
  Figure~\ref{fig:msddm_eyetest}. There is little evidence of trapping behavior
  or large hops. There are two reasons for this behavior. First, the width of
  hop length distributions are much smaller than those of the AD model. Closer
  examination of the characteristic MD trajectories shown in
  Figure~\ref{fig:solute_trajectories} reveal that hops tend to be an
  accumulation of a series of hops in the same direction. All of the hops in
  the MSDDM are negatively correlated which prevents this from happening. The
  second reason is a consequence of using a single hop length distribution for
  transitions. This was necessary because we could not collect enough data to
  fit all of the possible transition distributions and because we could not
  %MRS10: my feeling is (to keep in the back of the mind) that we may be able to add some sort of 
  %prior to fit the transition distributions. Lots of distributions can be fit with low amounts of data
  %if one makes enough assumptions about what the functional form of the distribution is. And we could
  %presumaly come up with something reasonable. But that's for another time. 
  correlate emissions coming from different hop distributions. Many
  transitions occur between two trapped states where the transitional hops are
  actually very small. Our model ignores this physical restriction which can
  cause the solute to drift rather than stay trapped.

@@

% LocalWords:  solute MSDs solutes solute's LLC BJC nanoporous maginn Lyotropic
% LocalWords:  micropollutants LLCs amphiphilic mesophases feng Bicontinuous EG
% LocalWords:  permeability hatakeyama zhou resel selectivities Edisonian CTRW
% LocalWords:  subdiffusive timescale timescales subdiffusion FBM RWF metzler
% LocalWords:  jeon neusius pdf Sokolov ony MSD MSMs chodera MSM AcOH URE MeOH
% LocalWords:  Hbonding hbonds hbond GROMACS berendsen gromacs der spoel hess
% LocalWords:  wt Na GA carboxylate Parrinello Rahman barostat rescale msd dt
% LocalWords:  superdiffusive EB sFBM min Hurwitz factoid MLEs autocovariance
% LocalWords:  evy sFLM FLM sFLMcut Dirichlet saiid fbm Stoev Taqqu MLE TODO SI
% LocalWords:  vvvv wehmeyer MSMbuilder pyEMMA featurization Phys Chem bayesian
% LocalWords:  Markovianity MSDDMs MSDDM resampling TBD MSD's CTRWs GCL acf ACH
% LocalWords:  urea's alkane overlayed SFLMcut underpredicted Gaussians HMM PRF
% LocalWords:  params msddm timeseries overpredicted GAANN ACS XSEDE ACI PSC QI
% LocalWords:  RMACC TOC Nanoscale Coscia lyotropic morphologies HII geens MFPT
% LocalWords:  darvishmanesh bouchaud integrator LINCS clauset sFBMcut pyemma
% LocalWords:  msmbuilder diffusivity Eq vt dz erfc permeabilities Kedem zoubi
% LocalWords:  Katchalsky's nanofiltration ultrafiltration trans convective py
% LocalWords:  bothmode todo eyetest Methanol's acid's subplots methanol's alt
% LocalWords:  timestep pre selectivtiy foulants subfigures RWFs roman github
% LocalWords:  al linkable nanopores

>>>>

  Up through our previous studies, our molecular models have provided valuable
  qualitative mechanistic insight with some quantitative support. This insight
  already allows us to speculate about new LLC monomer designs. However, they
  would be of greater value to a larger set of researchers if we could provide
  quantitative predictions of macroscopic observables such as solute flux and
  selectivity. Due to the size of the types of systems we are studying (at
  least 62,000 atoms) it is prohibitively expensive and time-consuming to run
  simulations longer than those performed for this work ($\sim$5 $\mu$s).  Even
  with the relatively large size of our system, we only simulate 24 solute
  trajectories per unit cell in order to minimize solute-solute interactions.
  This results in relatively high uncertainties in observables such as mean
  squared displacement (MSD), preventing reliable long timescale predictions.
  We are in need of a way of studying the collective motion of a much larger
  set of solutes over much longer timescales.
  %MRS11: Maybe this last sentence could be changed. Because you aren't really finding a way to study larger sets of 
  %solutes over larger timescales, you are using a limited amount of atomistic simulation into long time scale behavior efficiently. We can discuss variations as needed!

  %BJC: maybe this paragraph can be cut down?
  Mathematical descriptions of transport in complex separations membranes are a
  powerful way to understand mechanisms and formulate design principles.
  \cite{vinh-thang_predictive_2013,geens_transport_2006,darvishmanesh_mass_2016}
  %MRS11: this next sentence could probably be cut - it's kind of vague.
  The complexity of a well-fit model generally parallels the complexity of the
  transport mechanism being studied, as well as the transport information the
  model conveys. 
  In dense homogeneous membranes, the solution-diffusion model
  can extract diffusion and partition coefficients and has successfully
  predicted solute transport rates.~\cite{wijmans_solution-diffusion_1995}
  Analogously, pore-flow models yield predictions of diffusion coefficients and
  solute transport rates in nanoporous membranes.~\cite{paul_diffusive_1974}
  Modern single particle tracking approaches have taken researchers beyond
  continuum modeling allowing them to characterize complex diffusive
  behavior.~\cite{manzo_review_2015} At the molecular level, one can use
  molecular dynamics (MD) simulations to study both single particle dynamics
  and bulk transport properties with atomic-level
  insight.~\cite{coscia_chemically_2019,maginn_best_2018} All of these
  approaches facilitate generation of hypotheses about the molecular origins of
  separations by attempting to give a more intuitive understanding of how
  solutes move as a function of their environment, in turn suggesting
  experiments that could be performed.

  %MRS11: in this paragraph, the ``we can'' is vague, and doesn't
  %abscribe agency.  Is ``We'' our group? Is it the field as whole?
  %If it's something someone has done in the past, there should be
  %citations. If it's something that can be done in the future, the
  %phrasing should be a long the lines of ``it is possible to'', ``it
  %can be done'', etc.

@@

  %MRS11: should be clear if first model is previous publications, or what is presented here.
  %If what is presented here, should talk about how it will be presented, not that it was developed. 
  %MRS11: research activities happened in the past, presentation of research is described in the present.
  %or potentially future if describing what will be presented in the rest of the paper. 
  %BJC: replacement to above.
  We constructed our first model in order to characterize the hopping and
  trapping behavior displayed by solutes in our MD simulations. This type 
  of non-Brownian dynamical behavior results in subdiffusive transport which
  is well-described by anomalous diffusion (AD) theory.~\cite{metzler_random_2000,bouchaud_anomalous_1990,klages_anomalous_2008}

@@

  %MRS11: Introduction should not really give the answer so far.
  %Introduction should describe the questions that will bbe asked, but
  %not answer them.  That's not what is expected in an intro, it may confuse people (answers should be in the abstract).
  %MRS11: this next paragraph could use tweaks/cutting the most, but review the previous paragraphs as well.
  %MRS11: you can think about putting some of these results into conclusion, though - it reads well now, but might be benefitted from reemphasizing results if it's currently missing any of this.

@@

  Despite relatively good predictions of the MD MSD, qualitative mismatch
  between simulated MSDDM and MD trajectories suggest that the MSDDM may be
  getting the right answers for the wrong reasons. We plotted typical
  realizations of the MSDDM for each solute and compared them to MD in
  Figure~\ref{fig:msddm_eyetest}. There is little evidence of trapping behavior
  or large hops. There are two reasons for this behavior. First, the width of
  hop length distributions are much smaller than those of the AD model. Closer
  examination of the characteristic MD trajectories shown in
  Figure~\ref{fig:solute_trajectories} reveal that hops tend to be an
  accumulation of a series of hops in the same direction. All of the hops in
  the MSDDM are negatively correlated which prevents this from happening. The
  second reason is a consequence of using a single hop length distribution for
  transitions. This was necessary because we could not collect enough data to
  fit all of the possible transition distributions and because we could not
  %MRS10: my feeling is (to keep in the back of the mind) that we may be able to add some sort of 
  %prior to fit the transition distributions. Lots of distributions can be fit with low amounts of data
  %if one makes enough assumptions about what the functional form of the distribution is. And we could
  %presumably come up with something reasonable. But that's for another time. 
  correlate emissions coming from different hop distributions. Many
  transitions occur between two trapped states where the transitional hops are
  actually very small. Our model ignores this physical restriction which can
  cause the solute to drift rather than stay trapped.

@@

% LocalWords:  solute MSDs solutes solute's LLC BJC nanoporous maginn Lyotropic
% LocalWords:  micropollutants LLCs amphiphilic mesophases feng Bicontinuous EG
% LocalWords:  permeability hatakeyama zhou resel selectivities Edisonian CTRW
% LocalWords:  subdiffusive timescale timescales subdiffusion FBM RWF metzler
% LocalWords:  jeon neusius pdf Sokolov ony MSD MSMs chodera MSM AcOH URE MeOH
% LocalWords:  Hbonding hbonds hbond GROMACS berendsen gromacs der spoel hess
% LocalWords:  wt Na GA carboxylate Parrinello Rahman barostat rescale msd dt
% LocalWords:  superdiffusive EB sFBM min Hurwitz factoid MLEs autocovariance
% LocalWords:  evy sFLM FLM sFLMcut Dirichlet saiid fbm Stoev Taqqu MLE TODO SI
% LocalWords:  vvvv wehmeyer MSMbuilder pyEMMA featurization Phys Chem bayesian
% LocalWords:  Markovianity MSDDMs MSDDM resampling TBD MSD's CTRWs GCL acf ACH
% LocalWords:  urea's alkane overlayed SFLMcut underpredicted Gaussians HMM PRF
% LocalWords:  params msddm timeseries overpredicted GAANN ACS XSEDE ACI PSC QI
% LocalWords:  RMACC TOC Nanoscale Coscia lyotropic morphologies HII geens MFPT
% LocalWords:  darvishmanesh bouchaud integrator LINCS clauset sFBMcut pyemma
% LocalWords:  msmbuilder diffusivity Eq vt dz erfc permeabilities Kedem zoubi
% LocalWords:  Katchalsky's nanofiltration ultrafiltration trans convective py
% LocalWords:  bothmode todo eyetest Methanol's acid's subplots methanol's alt
% LocalWords:  timestep pre selectivtiy foulants subfigures RWFs roman github
% LocalWords:  al linkable nanopores klages
